### 2022-02-25 draft_paper.Rmd
add analyses and example rmd uses

---
title: "How do changes in ex vary between countries?"
author: "Jon Minton"
date: "22/02/2022"
output:
  word_document: default
  html_document: default
---

# Introduction

This notebook will compare annual changes in life expectancy at both birth ($e_0$) and at age 65 ($e_{65}$) in the UK and UK nations, with other select European nations, including a 'Synthetic Germany' created from a weighted average of East and West German data for common years. The method for creating this 'Synthetic Germany' is also described in this and linked documents.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Methods

## Data

Data were extracted from the [Human Mortality Database](https://mortality.org/) via the R package [`HMDHFDplus`](http://www.demogr.mpg.de/en/projects_publications/publications_1904/mpidr_technical_reports/reading_human_fertility_database_and_human_mortality_database_data_into_r_5438.htm).

The following countries are included in the comparison:

- England & Wales (`GBRTENW`)
- Scotland (`GBR_SCO`)
- UK as a whole (`GBR_UK`)
- France (`FRANTP`)
- Spain (`ESP`)
- Italy (`ITA`)
- Germany
   - Total Germany (`DEUTNP`)
   - East Germany (`DEUTE`)
   - West Germany (`DEUTW`)
   - Simulated/Synthetic Germany

```{r, echo=FALSE}
# load packages

pacman::p_load(here, tidyverse)

# load data

hmd_lt <- read_rds(here("data", "lifetables.rds"))

```

## Analysis

For each of these countries, and for males and females separately, we are interested in the annual changes in life expectancy at birth ($e_0$) and life expectancy at age 65 ($e_{65}$), from 1980 to the last available year for each country.

```{r, echo = FALSE}

countries_of_interest <- c(
  "GBRTENW",
  "GBR_SCO",
  "GBR_UK",
  "FRATNP",
  "ESP",
  "ITA",
  "DEUTNP",
  "DEUTE", 
  "DEUTW"
)

hmd_ex_selected_countries <- 
  hmd_lt %>% 
    filter(code %in% countries_of_interest) %>% 
    filter(Age %in% c(0, 65)) %>% 
    select(code, Year, Age, sex, ex) %>% 
    rename(x = Age, year = Year)

hmd_ex_selected_countries

```

# Results

## Descriptive Results /Starter Eyeballing

### Life expectancy trends

The following shows life expectancy at birth and at age 65 for selected nations:

Here it is at birth...

```{r, echo = FALSE, fig.cap = "Life Expectancy at birth"}

hmd_ex_selected_countries %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  ggplot(aes(x = year, y = ex, group = sex, colour = sex, linetype = sex)) + 
  geom_line() +
  facet_wrap(~code) +
  labs(
    x = "Year",
    y = "Life Expectancy at birth",
    title = "Life expectancies at birth for selected nations",
    subtitle = "1980 to 2020 or latest available year",
    caption = "Source: Human Mortality Database"
  )

```

Here it is at age 65

```{r, echo = FALSE, fig.cap = "Life Expectancy at age 65"}
hmd_ex_selected_countries %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  ggplot(aes(x = year, y = ex, group = sex, colour = sex, linetype = sex)) + 
  geom_line() +
  facet_wrap(~code) +
  labs(
    x = "Year",
    y = "Life Expectancy at age 65",
    title = "Life expectancies at age 65 for selected nations",
    subtitle = "1980 to 2020 or latest available year",
    caption = "Source: Human Mortality Database"
  )

```

Some insightful comment about what's going on.

Scotland looks pretty bad compared with the other countries.

## Change in life expectancy in these nations

```{r}
change_in_ex_selected_countries <- 
  hmd_ex_selected_countries %>% 
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
    ungroup()

change_in_ex_selected_countries
```

```{r}
change_in_ex_selected_countries %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(code~sex) + 
  geom_hline(yintercept = 0)

```

# Discussion

# Bibliography