### 2022-03-10 draft_graphs.Rmd
new doc with graphs for each country, age, sex and labels

---
title: "How do changes in ex vary between countries?"
author: "Jon Minton, Lucinda Hiam"
date: "22/02/2022"
output:
  html_document: default
  word_document: default
---

# To dos

- [ ] Create flextable by decade

# Introduction

Stalling of life expectancy improvements occurred in the UK from 2010 onwards. It was also seen in some other Western European countries. While the reason and severity of this is debated, some also contend the stalling of improvements seen is artefact following an unusual decade of improvement in 2000s (e.g. Murphy 2021).

- [ ] Introduction to be written once methods/results discussed and decided on

This notebook will compare annual changes in life expectancy at both birth ($e_0$) and at age 65 ($e_{65}$) from 1980 to 2019 (or the latest available year) in England and Wales, Scotland, France, Italy, Spains and Germany. Germany includes a 'Synthetic Germany' created from a weighted average of East and West German data for common years. The method for creating this 'Synthetic Germany' is covered in the methods section in brief and in Appendix X for more detail. In order to explore the hypothesis that the recent stalling in life expectancy, most marked in England and Wales, was an expected 'correction' after a decade of unusually fast life expectancy improvements, we have selected the years from 1980 onwards.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  tab.id = NULL,
  tab.cap = NULL,
  tab.topcaption = TRUE,
  tab.lp = "tab:"

  )
```

# Methods

## Data

Data were extracted from the [Human Mortality Database](https://mortality.org/) via the R package [`HMDHFDplus`](http://www.demogr.mpg.de/en/projects_publications/publications_1904/mpidr_technical_reports/reading_human_fertility_database_and_human_mortality_database_data_into_r_5438.htm).

The following countries are included in the comparison:

- England & Wales (`GBRTENW`)
- Scotland (`GBR_SCO`)
- UK as a whole (`GBR_UK`) #we haven't included UK but can do
- France (`FRANTP`)
- Spain (`ESP`)
- Italy (`ITA`)
- Germany
   - Total Germany (`DEUTNP`)
   - East Germany (`DEUTE`)
   - West Germany (`DEUTW`)
   - Simulated/Synthetic Germany

```{r, echo=FALSE}
# load packages

pacman::p_load(knitr, here, tidyverse, flextable) # using flextable instead of gt. It seems to be better 
# developed, more commonly used, and has more clearly labelled function names.

# load data

hmd_lt <- read_rds(here("data", "lifetables.rds"))

```

## Analysis

For each of these countries, and for males and females separately, we are interested in the annual changes in life expectancy at birth ($e_0$) and life expectancy at age 65 ($e_{65}$), from 1980 to the last available year for each country. In addition, we compare average life expectancy improvements over the 4 decades: 1980, 1990, 2000, and 2010.

We then compared whether the average between the decades was significant, using a regression model.

### Synthetic Germany

- To maybe be moved to the appendix later

To allow UK national trends to be compared with a single German population, we attempted to produce a 'Synthetic German' population with data for East and West Germany for years prior to reunification. We estimated that this 'Synthetic Germany' could be produced by using a weighted average of 20% East Germany, and 80% West German life expectancy trends. More precise estimates can be produced, and the methods used to reach this conclusion are detailed in Appendix X.

```{r, echo = FALSE}

countries_of_interest <- c(
  "GBRTENW",
  "GBR_SCO",
  "GBR_UK",
  "FRATNP",
  "ESP",
  "ITA",
  "DEUTNP",
  "DEUTE", 
  "DEUTW"
)

hmd_ex_selected_countries <- 
  hmd_lt %>% 
    filter(code %in% countries_of_interest) %>% 
    filter(Age %in% c(0, 65)) %>% 
    select(code, Year, Age, sex, ex) %>% 
    rename(x = Age, year = Year)

hmd_ex_selected_countries

# Code for producing synthetic Germany

source(here("R", "make_synthetic_germany_functions.R"))

series_east_f0 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTE") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  filter(sex == "female") %>% 
  arrange(year) %>% 
  pull(ex)

series_west_f0 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTW") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  filter(sex == "female") %>% 
  arrange(year) %>% 
  pull(ex)

tmp_df_f0 <- tibble(
  code = "DEUTSYNTH",
  year = 1980:2017,
  x = 0,
  sex = "female", 
  ex = make_synthetic_population(series_east_f0, series_west_f0, p_east = 0.2)
)

###
series_east_f65 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTE") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  filter(sex == "female") %>% 
  arrange(year) %>% 
  pull(ex)

series_west_f65 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTW") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  filter(sex == "female") %>% 
  arrange(year) %>% 
  pull(ex)

tmp_df_f65 <- tibble(
  code = "DEUTSYNTH",
  year = 1980:2017,
  x = 65,
  sex = "female", 
  ex = make_synthetic_population(series_east_f65, series_west_f65, p_east = 0.2)
)
##

series_east_m0 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTE") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  filter(sex == "male") %>% 
  arrange(year) %>% 
  pull(ex)

series_west_m0 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTW") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  filter(sex == "male") %>% 
  arrange(year) %>% 
  pull(ex)

tmp_df_m0 <- tibble(
  code = "DEUTSYNTH",
  year = 1980:2017,
  x = 0,
  sex = "male", 
  ex = make_synthetic_population(series_east_m0, series_west_m0, p_east = 0.2)
)

###

series_east_m65 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTE") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  filter(sex == "male") %>% 
  arrange(year) %>% 
  pull(ex)

series_west_m65 <- hmd_ex_selected_countries %>% 
  filter(code == "DEUTW") %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  filter(sex == "male") %>% 
  arrange(year) %>% 
  pull(ex)

tmp_df_m65 <- tibble(
  code = "DEUTSYNTH",
  year = 1980:2017,
  x = 65,
  sex = "male", 
  ex = make_synthetic_population(series_east_m65, series_west_m65, p_east = 0.2)
)

hmd_ex_selected_countries_with_synth <- bind_rows(
  hmd_ex_selected_countries,
  tmp_df_m0,
  tmp_df_m65,
  tmp_df_f0,
  tmp_df_f65
)

# Labels for codes

country_code_lookup <- 
  tribble(
    ~code, ~country,
    "DEUTNP", "Germany",
    "DEUTE", "East Germany",
    "DEUTW", "West Germany",
    "ESP", "Spain",
    "FRATNP", "France", 
    "ITA", "Italy",
    "GBRTENW", "England & Wales",
    "GBR_SCO", "Scotland",
    "DEUTSYNTH", "Synthetic Germany"
  )

```

# Results

## Descriptive Results

### Life expectancy trends

The following figures show life expectancy at birth and at age 65 for selected nations from 1980 to the latest available year for males and females.

```{r, echo = FALSE, fig.cap = "Life Expectancy at birth"}

hmd_ex_selected_countries_with_synth %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 0) %>% 
  ggplot(aes(x = year, y = ex, group = sex, colour = sex, linetype = sex)) + 
  geom_line() +
  facet_wrap(~country) +
  labs(
    x = "Year",
    y = "Life Expectancy at birth",
    title = "Life expectancies at birth for selected nations",
    subtitle = "1980 to 2020 or latest available year",
    caption = "Source: Human Mortality Database"
  )

```

```{r, echo = FALSE, fig.cap = "Life Expectancy at age 65"}
hmd_ex_selected_countries_with_synth %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  filter(x == 65) %>% 
  ggplot(aes(x = year, y = ex, group = sex, colour = sex, linetype = sex)) + 
  geom_line() +
  facet_wrap(~country) +
  labs(
    x = "Year",
    y = "Life Expectancy at age 65",
    title = "Life expectancies at age 65 for selected nations",
    subtitle = "1980 to 2020 or latest available year",
    caption = "Source: Human Mortality Database"
  )

```

- [ ] Some insightful comment about what's going on.

## Change in life expectancy in these nations

Next, we compare the annual change in life expectancy at birth and at aged 65 years for the nations from 1980 to the latest available year, by sex.

```{r, echo=FALSE}
change_in_ex_selected_countries <- 
  hmd_ex_selected_countries_with_synth %>% 
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
    ungroup()

change_in_ex_selected_countries
```

### Changes in life expectancy at birth

- [ ] make smoother lines coloured by sex
- [X] use consistent colours for all graphs showing separate sex results

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.height=8, fig.width=8}
change_in_ex_selected_countries %>% 
  filter(x == 0) %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex, colour=sex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(country~sex) + 
  geom_hline(yintercept = 0) + 
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(
    x = "Year",
    y = "Year-on-year change in life expectancy",
    title = "Annual change in life expectancy at birth, selected countries",
    subtitle = "Blue line: nonlinear smoother",
    caption =  "Source: Human Mortality Database. Synthetic Germany based on 20% East/80% West German population weighting"
  )

```

### Changes in life expectancy at age 65

- [ ] Use slightly stronger (and consistent) colours for sex smoother lines

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.height=8, fig.width=8}
change_in_ex_selected_countries %>% 
  filter(x == 65) %>% 
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  filter(between(year, 1980, 2020)) %>% 
  ggplot(aes(x = year, y = delta_ex, colour=sex)) + 
  geom_point(alpha = 0.2) + 
  stat_smooth() + 
  facet_grid(country~sex) + 
  geom_hline(yintercept = 0) + 
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(
    x = "Year",
    y = "Year-on-year change in life expectancy",
    title = "Annual change in life expectancy at age 65, selected countries",
    subtitle = "Blue line: nonlinear smoother",
    caption =  "Source: Human Mortality Database. Synthetic Germany based on 20% East/80% West German population weighting"
  )

```

## Time trends

Finally, we used time-trend analysis to explore the changes in life expectancy over the 4 decades: 1980s, 1990s, 2000s, and 2010s. This produces the average change for the decade, whether negative or positive, the standard error and its statistical significance for each country by sex.

- [X] Give this a title and further info (lower priority)
- [X] Look at splitting by starting age 
    - [ ] Look at further refining starting age split

```{r, echo = FALSE, warning = FALSE, message = FALSE}

hmd_ex_selected_countries_with_synth %>% 
    filter(year >= 1979) %>% 
    filter(!(code %in% c("DEUTE", "DEUTW", "DEUTNP"))) %>%  # Using only synthetic germany for longer time period
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
    nest() %>% 
    mutate(
      rwd_model = map(data, ~lm(delta_ex ~ 1, data = . )),
      ar_model  = map(data, ~lm(delta_ex ~ lag(delta_ex), data = .))
    ) %>% 
    mutate(
      aic_rwd = map_dbl(rwd_model, AIC),
      aic_ar  = map_dbl(ar_model, AIC)
    ) %>% 
    mutate(
      diff_aic = aic_ar - aic_rwd
    )  %>% 
  # filter(diff_aic > 0)
  mutate(
    tidied_ar_model = map(ar_model, broom::tidy)
  ) %>% 
  unnest(tidied_ar_model) %>% 
  filter(term == "(Intercept)") %>% 
  arrange(desc(estimate)) %>%
  left_join(country_code_lookup) %>% 
  mutate(country = factor(country, levels = c("England & Wales", "Scotland", "East Germany", "West Germany", "Germany", "Synthetic Germany", "Spain", "France", "Italy"))) %>% 
  select(code, country, x, sex, estimate, std.error) %>% 
  mutate(display_label = glue::glue("{country}_{x}_{sex}")) %>% 
  ggplot(aes(estimate, fct_reorder(display_label, estimate))) + 
  geom_point() +
  expand_limits( x = 0) + 
  geom_vline(xintercept = 0) + 
  labs(
    x = "Average annual life expectancy change",
    y = "Country-sex combination",
    title = "Average annual improvements by country and sex",
    subtitle = "Range: 1980 to 2020 or nearest available years",
    caption = "Source: Human Mortality Database"
  ) + 
  facet_wrap(~x)

```

### Average improvements by decade

- [ ] Add further introduction 
- [ ] Add statistical significance (though maybe this is a different analysis)

The table below shows average improvements by decade for England and Wales (standard error).

```{r, echo=FALSE, tab.cap = "Source: HMD"}

hmd_ex_selected_countries_with_synth %>% 
    filter(year >= 1979) %>% 
    group_by(code, x, sex) %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>%
    mutate(decade = year - (year %% 10)) %>% 
    filter(!is.na(delta_ex)) %>% 
  group_by(code, x, sex, decade) %>% 
  summarise(
    mean_delta_ex = mean(delta_ex),
    sd_delta_ex   = sd(delta_ex)
    ) %>% 
  ungroup() %>% 
  mutate(cell_contents = glue::glue("{round(mean_delta_ex, 3)} ({round(sd_delta_ex, 3)})")
  ) %>% 
  filter(code == "GBRTENW") %>% 
  filter(x == 0) %>% 
  select(sex, decade, cell_contents) %>% 
  spread(decade, cell_contents) %>% 
#  write.csv("clipboard")
  flextable() %>%
  # Want to make the header row bold
  vline(j = 1, part = "all") %>%
  set_caption(caption = "Average improvement in life expectancy by decade")

```

### Are the differences by decade statistically significant?

Are there significant differences in average rates of change in $e_x$ by decade? Another way of expressing this question is to ask whether knowing the decade in which an observation has been observed is informative as to what such values will be. This rephrased question was addressed by, for each starting age (0 or 65 years), sex and country, fitting two competing model specifications - $M_0$ and $M_A$ - and running an F test on the residuals of these two models. The Null model specification, $M_0$, regresses $\delta e_x$ (observed annual changes in $e_x$) against a single intercept term $\alpha$; put another way: $M_0$ assumes that all observations are drawn from a Normal distribution with a mean value of $\alpha$, and that this is consistent across all observed decades. The alternative model specification, $M_A$, also includes the $\alpha$ term, but also includes dummy variables to indicate which decade the realised value of $\delta e_0$ corresponds to. (Because one of the decades needs to be the reference category against which other decades are compared, there are one fewer decadal dummy variables than decades in the dataset for a given country, and the $\alpha$ parameter in $M_A$ estimates the mean change in $\delta e_0$ in the first decade, and so cannot be compared directly with the $\alpha$ parameter in $M_0$.) A P-value from the F test comparing $M_0$ and $M_A$ of less than 0.05 is taken to indicate that $M_A$ should be preferred to $M_0$, and so that there are systemic differences in average rates of improvement by decade for a given country, sex, and starting age.

We found the only significant change decade-on-decades was for East Germany for females at birth and at aged 65 years. This is consistent with existing literature in this area, showing the convergence of life expectancy for females in East and West Germany following reunification.

```{r, cache=TRUE}

dta_for_decade_comparison <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "GBRTENW") %>% 
      filter(year >= 1979) %>% 
      # ggplot(aes(x = year, y = ex)) +
      #   geom_line()
      arrange(year) %>% 
      mutate(delta_ex = ex - lag(ex)) %>% 
    #   ggplot(aes(x = year, y = delta_ex)) + 
    # geom_point() + geom_line() +
    # geom_hline(yintercept = 0)
      mutate(
        decade = year - (year %% 10)
      ) %>% 
  filter(!is.na(delta_ex))

model_null <- 
  lm(delta_ex ~ 1, data = dta_for_decade_comparison)

model_alt <- 
  lm(delta_ex ~ factor(decade), data= dta_for_decade_comparison)

summary(model_null)

summary(model_alt)

anova(model_null, model_alt)

# to get the value of interest do tmp[2,6]

# Generalising

tidy_dataset <- function(df){
  df %>% 
      filter(year >= 1979) %>% 
      arrange(year) %>% 
      mutate(delta_ex = ex - lag(ex)) %>% 
      mutate(
        decade = year - (year %% 10)
      ) %>% 
  filter(!is.na(delta_ex))
}

make_null_model <- function(df){
  lm(delta_ex ~ 1, data = df)
}

all_null_alt_models <-
  hmd_ex_selected_countries_with_synth %>%
    group_by(x, sex, code) %>% 
    nest() %>% 
    mutate(
      prep_data = map(data, tidy_dataset) 
    ) %>% 
    mutate(
      model_null = map(prep_data, ~lm(delta_ex ~ 1, data = .)),
      model_alt  = map(prep_data, ~lm(delta_ex ~ factor(decade), data = .))
  #    model_null = map(prep_data, make_null_model)
    ) %>% 
    mutate(
      anova_results = map2(model_null, model_alt, anova)
    ) %>% 
    mutate(
      anova_pval = map_dbl(anova_results, ~.[2,6])
    )

all_null_alt_models %>% 
  filter(anova_pval < 0.05) -> tmp

summary(tmp$model_alt[[1]])

summary(tmp$model_alt[[2]])

  # build null model
  # build alt model
  # do anova
  # get p value from anova

```

### Joinpoint/Segmented regression approach

Though there may be systemic differences in average changes in life expectancy over time, such differences may not be patterned by decade, and so the above model specification may not have been appropriate to detect any such changes. Instead of looking for systemic differences by decade, we can instead look at whether there are any systemic and detectable 'breaks' in the trends over time. If our data were life expectancy over time, this could be achieved through joinpoint analysis, in which the slope is allowed to change over time. As instead are data are annual changes in life expectancy, we are interested in detecting changes in intercept over time. If we define our Null (no change) model as follows:

$$
M_0 : y(t) = \alpha + e
$$
Where $e \backsim N(0, \sigma^2)$. i.e. all observations $y$ are draws from $N(\alpha, \sigma^2)$. The alternative model specification against which this is compared is then:

$$
M_A : y(t) = \alpha_0 + \beta_1 T(t) + e 
$$
Where $T(t)$ is defined as $0$ where $t \lt \tau$ and $1$ where $t \geq \tau$. ($\alpha_0$ is used in $M_A$ to indicate that this parameter is not directly comparable with $\alpha$ in $M_0$).

This specification shows that there is an additional parameter $\tau$ to be determined in order to produce $M_A$. This $\tau$ parameter is the breakpoint in the series. The selection of $\tau$ will be determined by numeric optimisation, by selecting the value of $\tau$ that minimised AIC.

Once the best possible $M_A$, using the best value of $\tau$, is selected, $M_A$ will be compared with $M_0$ using an F-test, allowing us to determine whether there is convincing evidence of any specific break in the data.

We will start with a single country, starting age, and sex, and with manual selection of $\tau$, then generalise and automate further.

## Graphs of annual change in life expectancy

The below graphs show the annual change in life expectancy at birth and at aged 65 years for males and females for the nations examined. The pink and green lines are non-linear smoothed.

NB: those with statistically significant 'breakpoints' in annual change in life expectancy are:
- Germany (need to re-download from GIT)

### England and Wales

```{r, echo = FALSE, warning=FALSE, message = FALSE}

#ENW female at birth
f_e0_ENW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "GBRTENW") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_ENW <- f_e0_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ENW male at birth
m_e0_ENW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "GBRTENW") %>% 
      filter(year >= 1979)

  m_e0_ENW <- m_e0_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #ENW female at age 65
f_e65_ENW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "GBRTENW") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_ENW <- f_e65_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ENW male at aged 65
m_e65_ENW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "GBRTENW") %>% 
      filter(year >= 1979)

 m_e65_ENW <- m_e65_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

 #add labels
f_e0_ENW + labs(y = "Annual change in e0", title = "England and Wales", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")

m_e0_ENW + labs(y = "Annual change in e0", title = "England and Wales", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")

f_e65_ENW + labs(y = "Annual change in e65", title = "England and Wales", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")

m_e65_ENW + labs(y = "Annual change in e65", title = "England and Wales", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

  #to make grid, first assign to object
# f_e0_ENW <- f_e0_ENW + labs(y = "Annual change in e0", title = "England and Wales", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
# 
# m_e0_ENW <- m_e0_ENW + labs(y = "Annual change in e0", title = "England and Wales", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
# 
# f_e65_ENW <- f_e65_ENW + labs(y = "Annual change in e65", title = "England and Wales", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
# 
# m_e65_ENW <- m_e65_ENW + labs(y = "Annual change in e65", title = "England and Wales", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

#changing labels for grid to work
# f_e0_ENW <- f_e0_ENW + labs(y = "Annual change in e0", title = "England and Wales", subtitle = "Life expectancy changes at birth for females")
# m_e0_ENW <- m_e0_ENW + labs(y = "Annual change in e0", subtitle = "Life expectancy changes at birth for males")
# f_e65_ENW <- f_e65_ENW + labs(y = "Annual change in e65")
# m_e65_ENW <- m_e65_ENW + labs(y = "Annual change in e65", caption = "Source: Human Mortality Database")
# #put side by side, needs package gridExtra then uses grid.arrange(plot1, plot2, ncol=2)
# grid.arrange(f_e0_ENW, m_e0_ENW, f_e65_ENW, m_e65_ENW, ncol=2, nrow=2)
# #grid.arrange(f_e65_ENW, m_e65_ENW, ncol=2, nrow=2)

```

### Scotland
```{r, echo = FALSE, warning=FALSE, message = FALSE}
#SCO female at birth
f_e0_SCO <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "GBR_SCO") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_SCO <- f_e0_SCO %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #SCO male at birth
m_e0_SCO <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "GBR_SCO") %>% 
      filter(year >= 1979)

  m_e0_SCO <- m_e0_SCO %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #SCO female at age 65
f_e65_SCO <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "GBR_SCO") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_SCO <- f_e65_SCO %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #SCO male at aged 65
m_e65_SCO <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "GBR_SCO") %>% 
      filter(year >= 1979)

 m_e65_SCO <-  m_e65_SCO %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### France

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#FRA female at birth
f_e0_FRA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "FRATNP") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_FRA <- f_e0_FRA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #FRA male at birth
m_e0_FRA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "FRATNP") %>% 
      filter(year >= 1979)

  m_e0_FRA <- m_e0_FRA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #FRA female at age 65
f_e65_FRA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "FRATNP") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_FRA <- f_e65_FRA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #FRA male at aged 65
m_e65_FRA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "FRATNP") %>% 
      filter(year >= 1979)

 m_e65_FRA <-  m_e65_FRA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### Spain

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#ESP female at birth
f_e0_ESP <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "ESP") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_ESP <- f_e0_ESP %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ESP male at birth
m_e0_ESP <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "ESP") %>% 
      filter(year >= 1979)

  m_e0_ESP <- m_e0_ESP %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #ESP female at age 65
f_e65_ESP <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "ESP") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_ESP <- f_e65_ESP %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ESP male at aged 65
m_e65_ESP <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "ESP") %>% 
      filter(year >= 1979)

 m_e65_ESP <-  m_e65_ESP %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### Italy

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#ITA female at birth
f_e0_ITA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "ITA") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_ITA <- f_e0_ITA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ITA male at birth
m_e0_ITA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "ITA") %>% 
      filter(year >= 1979)

  m_e0_ITA <- m_e0_ITA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #ITA female at age 65
f_e65_ITA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "ITA") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_ITA <- f_e65_ITA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #ITA male at aged 65
m_e65_ITA <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "ITA") %>% 
      filter(year >= 1979)

 m_e65_ITA <-  m_e65_ITA %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   #Trying to add legend
   #stat_smooth(aes(colour = "aquamarine4"), show.legend = TRUE, labels = "Smoothed")
  stat_smooth(colour = "aquamarine4")

  #add labels
f_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### Germany: Synthetic

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#DEUTS female at birth
f_e0_DEUTS <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTSYNTH") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_DEUTS <- f_e0_DEUTS %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTS male at birth
m_e0_DEUTS <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTSYNTH") %>% 
      filter(year >= 1979)

  m_e0_DEUTS <- m_e0_DEUTS %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #DEUTS female at age 65
f_e65_DEUTS <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTSYNTH") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_DEUTS <- f_e65_DEUTS %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTS male at aged 65
m_e65_DEUTS <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTSYNTH") %>% 
      filter(year >= 1979)

 m_e65_DEUTS <-  m_e65_DEUTS %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### Germany: East

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#DEUTE female at birth
f_e0_DEUTE <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTE") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_DEUTE <- f_e0_DEUTE %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTE male at birth
m_e0_DEUTE <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTE") %>% 
      filter(year >= 1979)

  m_e0_DEUTE <- m_e0_DEUTE %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #DEUTE female at age 65
f_e65_DEUTE <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTE") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_DEUTE <- f_e65_DEUTE %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTE male at aged 65
m_e65_DEUTE <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTE") %>% 
      filter(year >= 1979)

 m_e65_DEUTE <-  m_e65_DEUTE %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

### Germany: West

```{r, echo = FALSE, warning=FALSE, message = FALSE}
#DEUTW female at birth
f_e0_DEUTW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTW") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e0_DEUTW <- f_e0_DEUTW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTW male at birth
m_e0_DEUTW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 0) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTW") %>% 
      filter(year >= 1979)

  m_e0_DEUTW <- m_e0_DEUTW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4")

  #DEUTW female at age 65
f_e65_DEUTW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "female") %>% 
      filter(code == "DEUTW") %>% 
      filter(year >= 1979)

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_DEUTW <- f_e65_DEUTW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

  #DEUTW male at aged 65
m_e65_DEUTW <- 
  hmd_ex_selected_countries_with_synth %>% 
      filter(x == 65) %>% 
      filter(sex == "male") %>% 
      filter(code == "DEUTW") %>% 
      filter(year >= 1979)

 m_e65_DEUTW <-  m_e65_DEUTW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "aquamarine4", show.legend = TRUE)

  #add labels
f_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

# Discussion

- **Limitations**: fewer years for comparison in 2010s for some nations; comparing different time periods will provide different results e.g. New Labour years in E&W.

# Appendix A: Construction of Synthetic Germany

- Add further details here.

# Bibliography