### 2022-06-02 notebooks/old_bp_cp_comparison.Rmd

---
title: "R Notebook"
output: html_notebook
---

# Code and notes on breakpoint and changepoint

Below is a chunk of code trying to make sense of the changepoint and breakpoint analyses when only comparing one changepoint/breakpoint against zero changepoints/breakpoints

# Start

```{r eval = FALSE}
# We can now run the above function, which creates a model, for a wide range of possible tau values
# This is the sequence listed and named tau_candidate, which is every whole number from 
# 1982 to 2015
# For each of these models, we can then get a measure of model fit. We're going to use AIC, though 
# could have used other measures. With AIC (and with BIC, and RMSE etc) better fit is indicated with 
# a lower value

source(here("R", "breakpoint_functions.R"))

# Let's see if this now works in the functional programming context, using 
# nest and tibble elements 
# hmd_ex_selected_countries_with_synth %>%
#   filter(year >= 1979) %>% 
#   group_by(code, x, sex) %>% 
#   nest() %>% 
#   mutate(
#     best_tau = map_dbl(data, get_best_tau_from_optim)
#   )
# I think this is getting stuck on local optima, so instead maybe 
# the more exhaustive search should be used after all! 
# There are two alternative approaches: 
# 1) Use an exhaustive grid search approach 
# 2) Use an algorithm which is less likely to get stuck in local optima
# The first approach is much more sensible, but the second might be a cool
# stretch goal as well! 
# let's use a heavily modified version of the above function to do the grid search

compare_best_breakmodel_with_null <- 
  hmd_ex_selected_countries_with_synth %>%
    filter(year >= 1979) %>% 
    group_by(code, x, sex) %>% 
    nest() %>% 
    mutate(
      best_tau = map_dbl(data, get_best_tau_from_gridsearch),
      all_taus = map(data, get_best_tau_from_gridsearch, what = "all")
    ) %>% 
    mutate(
      model_null = map(
        data, 
        ~ .x %>% 
          arrange(year) %>% 
          mutate(delta_ex = ex - lag(ex)) %>% 
          filter(!is.na(delta_ex)) %>% 
          lm(delta_ex ~ 1, data = .)
      )
    ) %>% 
    mutate(
      model_alt = map2(
        best_tau, data,
      ~ .y %>% 
        arrange(year) %>% 
        mutate(delta_ex = ex - lag(ex)) %>% 
        filter(!is.na(delta_ex)) %>% 
        mutate(T_param = ifelse(year < .x, FALSE, TRUE)) %>% 
        lm(delta_ex ~ T_param, data = .)
      )
    ) %>% 
  mutate(
    aic_alt  = map_dbl(model_alt, AIC),
    aic_null = map_dbl(model_null, AIC)
  ) %>% 
  mutate(
    alt_preferred = map2_lgl(aic_alt, aic_null, ~ifelse(.x < .y, TRUE, FALSE))
  ) %>% 
  mutate(
    anova_results = map2(model_alt, model_null, anova)
  ) %>% 
  mutate(
    ftest_pval = map_dbl(anova_results, ~.[2,6]),
    stat_sig_preference = ftest_pval < 0.05
  )

```

Let's now try to summarise the above results intelligently.

We want to know:

- [ ] Best year
- [ ] Whether stat sig over null 
- [ ] Estimate before break
- [ ] Estimate after break

Let's take the object above then get this info.

```{r}
compare_best_breakmodel_with_null %>% 
  mutate(population = glue::glue("{code}_{sex}_{x}")) %>% 
  ggplot(aes(best_tau, fct_reorder(population, best_tau), colour = stat_sig_preference)) + 
  geom_point()

```

I'm now a bit concerned by how many picked 2015. Is this an artefact of when the data finished?

To test this further I've now saved AIC by proposed year so let's look at how clear a 'winner' is for each population

```{r}
compare_best_breakmodel_with_null %>% 
  select(code, x, sex, best_tau, all_taus, ftest_pval) %>% 
  unnest(all_taus) %>% 
  ungroup() %>% 
  mutate(population = glue::glue("{code}_{x}_{sex}")) %>% 
  mutate(
    ftest_cat = case_when(
      ftest_pval < 0.01 ~ "<0.01",
      between(ftest_pval, 0.01, 0.05) ~ "0.01-0.05",
      between(ftest_pval, 0.05, 0.10) ~ "0.05-0.10",
      TRUE                            ~ "Not Stat. Sig."
    ),
    ftest_cat = factor(ftest_cat, 
      levels = c("<0.01", "0.01-0.05", "0.05-0.10", "Not Stat. Sig."),
      ordered = TRUE
    )
  ) %>% 
  ggplot(aes(tau, aic)) +
  geom_line() + geom_point(aes(colour = ftest_cat, alpha = tau == best_tau)) + 
  facet_wrap(~population, scales = "free_y") +
  scale_colour_manual(
    values = c(
      "<0.01" = "darkred", 
      "0.01-0.05"  = "red", 
      "0.05-0.10" = "orange", 
      "Not Stat. Sig." = "lightblue"
      )
  ) + 
  scale_alpha_manual(values = c(`FALSE` = 0, `TRUE` = 1), guide = "none") +
  labs(
    title = "Penalised fit (AIC) against proposed changepoint year, various population",
    subtitle = "Best fit identified indicate with a point, whose colour indicates statistical significance of changepoint model compared with Null model",
    colour = "Statistical\nsignificance", 
    x = "Year proposed",
    y = "AIC (lower is better)"
  )
```

I think this is a fairly complex visualisation, but hopefully usable either in the main paper or (more likely) supplementary material.

Here's a table attempting to summarise (subjectively) the above graphs:

| **Population**    | **x** | **Sex** | **Main Breakpoint (Stat sig.)** | **Other possible breakpoints** |
|-------------------|-------|---------|---------------------------------|--------------------------------|
| East Germany      | 0     | male    | 1991 **                         | No                             |
| East Germany      | 0     | female  | 2005 **                         | 2015                           |
| East Germany      | 65    | male    | 1991 **                         | No                             |
| East Germany      | 65    | female  | 1991 **                         | No                             |
| Germany           | 0     | male    | 2015 **                         | 2007ish                        |
| Germany           | 0     | female  | 2007 **                         | 2005-2008                      |
| Germany           | 65    | male    | 2015 **                         | No                             |
| Germany           | 65    | female  | 2015 *                          | 2005-2008                      |
| Synthetic Germany | 0     | male    | 2015 **                         | Weak 2007                      |
| Synthetic Germany | 0     | female  | 2007 **                         | 2005-2008, 2015                |
| Synthetic Germany | 65    | male    | 2015 **                         | No                             |
| Synthetic Germany | 65    | female  | 2015 *                          | Weak 2005-2007                 |
| West Germany      | 0     | male    | 2015 **                         | Weak 2007                      |
| West Germany      | 0     | female  | 1988 **                         | 2005-2008ish                   |
| West Germany      | 65    | male    | 2015 **                         | No                             |
| West Germany      | 65    | female  | 2015 *                          | 2008 ish (Weak)                |
| Spain             | 0     | male    | 1983 *                          | No                             |
| Spain             | 0     | female  | 1983 *                          | No                             |
| Spain             | 65    | male    | 2000 (NS)                       | 2015                           |
| Spain             | 65    | female  | 2014 (NS)                       | 1983?                          |
| France            | 0     | male    | 2015 (NS)                       | ?                              |
| France            | 0     | female  | 2015 (NS)                       | ?                              |
| France            | 65    | male    | 2015 **                         | No                             |
| France            | 65    | female  | 2015 *                          | No                             |
| Scotland          | 0     | male    | 2015 ***                        | No                             |
| Scotland          | 0     | female  | 1980 *                          | 2015 (Strong)                  |
| Scotland          | 65    | male    | 1994 (NS)                       | No?                            |
| Scotland          | 65    | female  | 1981 (NS)                       | No                             |
| England & Wales   | 0     | male    | 2013 ***                        | 2012-2015ish                   |
| England & Wales   | 0     | female  | 2012 **                         | Around 2015                    |
| England & Wales   | 65    | male    | 2012 *                          | 2015                           |
| England & Wales   | 65    | female  | 2012 (NS)                       | 2015                           |
| Italy             | 0     | male    | 1981 (NS)                       | 2015                           |
| Italy             | 0     | female  | 2005 (NS)                       | Weak 2015                      |
| Italy             | 65    | male    | 1984 (NS)                       | 1981                           |
| Italy             | 65    | female  | 2005 (NS)                       | 1984                           |

Let's now try to double-check the above results by running breakpoint analysis using the `segmented` package

```{r}
source(here("R", "segmented_script.R"))

```

Now I'm going to plot the breakpoints + SEs on the data itself

```{r}
segmented_breakpoint_models %>% 
  unnest(data) %>% 
  ggplot(aes(x = year, y = ex, colour = sex)) + 
  geom_line() + 
  facet_grid(x ~ code, scales = "free_y") + 
  geom_vline(aes(xintercept = breakpoint, color = sex))

```

I'm now going to compare the breakpoints from the segmented package with the best point from the line/gridsearch approach

```{r}
compare_best_breakmodel_with_null %>% 
  select(code, x, sex, best_tau, all_taus, ftest_pval) %>% 
  unnest(all_taus) %>% 
  ungroup() %>% left_join(
    segmented_breakpoint_models %>% 
      select(code, x, sex, bp_ftest_pval = ftest_pval, bp_breakpoint = breakpoint, bp_breakpoint_se = breakpoint_se)
  ) %>% 
  mutate(population = glue::glue("{code}_{x}_{sex}")) %>% 
  mutate(
    ftest_cat = case_when(
      ftest_pval < 0.01 ~ "<0.01",
      between(ftest_pval, 0.01, 0.05) ~ "0.01-0.05",
      between(ftest_pval, 0.05, 0.10) ~ "0.05-0.10",
      TRUE                            ~ "Not Stat. Sig."
    ),
    ftest_cat = factor(ftest_cat, 
      levels = c("<0.01", "0.01-0.05", "0.05-0.10", "Not Stat. Sig."),
      ordered = TRUE
    )
  ) %>% 
  mutate(
    bp_ftest_cat = case_when(
      bp_ftest_pval < 0.01 ~ "<0.01",
      between(bp_ftest_pval, 0.01, 0.05) ~ "0.01-0.05",
      between(bp_ftest_pval, 0.05, 0.10) ~ "0.05-0.10",
      TRUE                            ~ "Not Stat. Sig."
    ),
    bp_ftest_cat = factor(bp_ftest_cat, 
      levels = c("<0.01", "0.01-0.05", "0.05-0.10", "Not Stat. Sig."),
      ordered = TRUE
    )
  ) %>% 
  ggplot(aes(tau, aic)) +
  geom_line() + 
  geom_point(aes(colour = ftest_cat, alpha = tau == best_tau)) + 
  facet_wrap(~population, scales = "free_y") +
  scale_colour_manual(
    values = c(
      "<0.01" = "darkred", 
      "0.01-0.05"  = "red", 
      "0.05-0.10" = "orange", 
      "Not Stat. Sig." = "lightblue"
      )
  ) + 
  geom_vline(
    aes(
      xintercept = bp_breakpoint,
      colour = bp_ftest_cat  
    )
  ) + 
  geom_rect(
    aes(
      xmin = bp_breakpoint - 1.96 * bp_breakpoint_se, 
      xmax = bp_breakpoint + 1.96 * bp_breakpoint_se
    ),
      ymin = -Inf, ymax = Inf,
    alpha = 0.01
  ) + 
  scale_alpha_manual(values = c(`FALSE` = 0, `TRUE` = 1), guide = "none") +
  labs(
    title = "Penalised fit (AIC) against proposed changepoint year, various population",
    subtitle = "Best fit identified indicate with a point, whose colour indicates statistical significance of changepoint model compared with Null model",
    colour = "Statistical\nsignificance", 
    x = "Year proposed",
    y = "AIC (lower is better)",
    caption = "Vertical lines and bands show estimates and 95% CIs from segmented package"
  )

ggsave("figures/aic_from_gridsearch_and_segmented_bp_estimates.png", height = 20, width = 30, units = "cm", dpi = 150)
```

And now for change in ex

- [X] Jon to figure out why this doesn't work...
    - I think it doesn't work because the `segmented` package functions expect the line to be joined up. Instead it looks like the [structchange](https://cran.r-project.org/web/packages/strucchange/index.html) package can be used for more general model specifications.